---
title: "Jurisprudence and Social Policy - Math Camp 2016"
author: "Aniket Kesari"
date: "August 3, 2016"
output:
  beamer_presentation: null
  incremental: no
  ioslides_presentation: default
  latex_engine: lualatex
  slidy_presentation: default
subtitle: Probability Demo - Hot Hand
---

## The Hot Hand

- Probabilists debate whether the 'hot hand' is a real phenomenon
- The 'hot hand' refers to the idea that if a person experiences success as a result of a probabilistic event, the next attempt has a higher probability of success
- Tversky, Gilovich, Vallone (1985) disputed the 'hot hand theory' for college basketball players at Stanford and Cornell
    - Attributed the fallacy to low numerical literacy among the general public and sports commentators
    - Reflected a failure in understanding of probabilistic events and long-run frequencies
- However, more recent research has indicated that the 'hot hand' may indeed exist in certain applications

## Kobe's Hot Hand

- Kobe Bryant earned the MVP award for the 2009 NBA Finals when he led the LA Lakers to victory over the Orlando Magic
- At the time, many sports commentators remarked on his 'hot hand'.  Were they right?
- Let's look at some data and find out:

```{r echo = TRUE}
download.file("http://www.openintro.org/stat/data/kobe.RData", destfile = "kobe.RData")
load("kobe.RData")
```

## Kobe's Hot Hand

- The data show us all of Kobe's attempted shots during the course of the five-game (best of seven) NBA Finals
- In the 'basket' column, a 'H' indicates that Kobe made the shot, a 'M' indicates that he missed
- A streak is defined as the number of H's that occur before a M
- For example:
```{r echo=FALSE}
print(kobe$basket[1:9])
```
- In the first 9 rows, we have six streaks of lengths 1, 0, 2, 0, 0, and 0 respectively

## Kobe's Hot Hand

- Using the custom function 'calc_streak' (built in when you loaded the data, otherwise ask Aniket how to define the function) to calculate the distribution of Kobe's streak
- Then plot it to get a sense of its shape

## Kobe's Hot Hand

```{r echo=TRUE}
kobe_streak <- calc_streak(kobe$basket)
barplot(table(kobe_streak), xlab="Streak Length", ylab = "Frequency", main="Barplot of Kobe's Streaks in 2009 NBA Finals")
```

## Independence

- There are some longer streaks here, but how can we tell if there's a hot hand at work?
- *Independence* is key to understanding this
    - Two events are independent if the outcome of one does not affect the probability of the outcome of a second event
- In our example, a hot hand would mean that making a basket affects the probability of making the second basket

**Hot Hand**
$$
P(shot1=H)=.45
$$

$$
P(shot2=H|shot1=H)=.6
$$

**No Hot Hand**
$$
P(shot1=H)=.45
$$

$$
P(shot2=H|shot1=H)=.45
$$

## Independence

- Before turning to Kobe, let's look at an example of independence that we know works, coin flips
- Try this code:

```{r echo=TRUE}
outcomes <- c(0,1) #define what we're sampling from
fair_coin <- sample(outcomes, size=10000, replace=TRUE) #sample 10,000 times with replacement drawing from all of the choices with equal probability
S <- cumsum(fair_coin) #Add up all of the successes
r <- S/(1:length(fair_coin)) #Divide the successes by the overall number of trials
```

## Independence

- Now let's plot it (what happens as n gets bigger?):

```{r echo=TRUE}
plot(r, ylim=c(0,1),type="l")
```

## Independent Basketball Shooter

- OK, so we see that as 'n' gets larger, the long-run frequency converges to the independent probability (.5 in the case of a coin)
- Now let's do this with a hypothetical independent shooter who makes 45% of her shots:

```{r echo=TRUE}
outcomes <- c("H","M")
sim_basket <- sample(outcomes, size=133, replace=TRUE, prob=c(.45,.55))
```

## Problems

```{r echo=TRUE}
indpt_shooter_streak <- calc_streak(sim_basket)
indpt_shooter_streak <- data.frame(indpt_shooter_streak)
```

## Problems

```{r echo=TRUE}
par(mfrow=c(1,2))
barplot(table(indpt_shooter_streak), xlab="Streak Length", ylab= "Frequency", main="Barplot of Independent Shooter's Streaks in 2009 NBA Finals")
barplot(table(kobe_streak), xlab="Streak Length", ylab = "Frequency", main="Barplot of Kobe's Streaks in 2009 NBA Finals")
```